### 2025-03-09 base-rates.qmd
20250308 - citations

The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
For instance, the base rate of HIV is the probability of developing HIV.
In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).

@@

Base rates are also relevant to fantasy football.
Unlike yardage (e.g., passing yards, rushing yards, receiving yards), touchdowns occur *relatively* less frequently.
Whereas a solid Wide Receiver may log 100+ receptions and 1,200+ yards in a season, they may have "only" 8–14 receiving touchdowns in a given season.
As noted in @sec-differentErrorsDifferentCosts, lower base-rate events—including touchdowns—are harder to predict accurately.
As noted by [Harris (2012)](https://www.espn.com/fantasy/football/ffl/story?page=nfldk2k12_vbdwork): "NFL statistical projections are basically impossible to get right. (Take it from someone who helps create them for a living.) Yes, we can do a passable job with yardage totals for players who don't suffer unexpected injuries or depth-chart pratfalls. But so much of fantasy football hinges on touchdowns, and touchdowns are impossibly difficult to predict from season to season (let alone week to week)." (archived at <https://perma.cc/4QNH-J2LD>).
Thus, it is important not to lend too much credence to predictions of touchdowns.
Focus on other things that may be more predictable (and that may be indirectly prognostic of touchdowns) such as yards, carries/targets, receptions, depth of targets, red zone carries/targets, short distance carries/targets, etc.

@@

An interactive probability nomogram is available at the following link: <https://jamaevidence.mhmedical.com/data/calculators/LR_nomogram.html> (archived at <https://perma.cc/Z3SW-QMJ3>).

@@

A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):

>>>>

The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
For instance, the base rate of HIV is the probability of developing HIV.
In the U.S., the prevalence rate of HIV is ~0.4% of the adult population [@AIDSVu2022; archived at <https://perma.cc/8GE6-GAPC>].

@@

Base rates are also relevant to fantasy football.
Unlike yardage (e.g., passing yards, rushing yards, receiving yards), touchdowns occur *relatively* less frequently.
Whereas a solid Wide Receiver may log 100+ receptions and 1,200+ yards in a season, they may have "only" 8–14 receiving touchdowns in a given season.
As noted in @sec-differentErrorsDifferentCosts, lower base-rate events—including touchdowns—are harder to predict accurately.
As noted by @Harris2012: "NFL statistical projections are basically impossible to get right. (Take it from someone who helps create them for a living.) Yes, we can do a passable job with yardage totals for players who don't suffer unexpected injuries or depth-chart pratfalls. But so much of fantasy football hinges on touchdowns, and touchdowns are impossibly difficult to predict from season to season (let alone week to week)." (archived at <https://perma.cc/4QNH-J2LD>).
Thus, it is important not to lend too much credence to predictions of touchdowns.
Focus on other things that may be more predictable (and that may be indirectly prognostic of touchdowns) such as yards, carries/targets, receptions, depth of targets, red zone carries/targets, short distance carries/targets, etc.

@@

An interactive probability nomogram is available from @Altarejos2025 at the following link: <https://jamaevidence.mhmedical.com/data/calculators/LR_nomogram.html> (archived at <https://perma.cc/Z3SW-QMJ3>).

@@

A probability nomogram calculator is available from @Schwartz2006 at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:

### 2025-03-09 basic-statistics.qmd
20250308 - R packages
20250308 - citations

The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.

@@

A plot of statistical power is in @fig-nhst.

@@

- <https://rpsychologist.com/pvalue/> (archived at <https://perma.cc/JP9F-9ZVY>)
- <https://rpsychologist.com/d3/pdist/> (archived at <https://perma.cc/BE96-8LSJ>)
- <https://rpsychologist.com/d3/nhst/> (archived at <https://perma.cc/ZU9A-37F3>)

@@

For an online, interactive statistical decision tree to help you decide which statistical analysis to use, see here: <https://www.statsflowchart.co.uk>

@@

Knowing any three of the following, you can calculate the fourth: statistical power, [effect size](#sec-practicalSignificance), [sample size](#sec-sampleVsPopulation), and [alpha level](#sec-statisticalSignificance).
Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
For free point-and-click software for calculating statistical power, see G\*Power: <https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower.html>

@@

- <https://aguinis.shinyapps.io/ml_power/>
- <https://koumurayama.shinyapps.io/tmethod_mlm/>
- <https://webpower.psychstat.org/wiki/models/index>

@@

- <https://sjak.shinyapps.io/power4SEM/>
- <https://sempower.shinyapps.io/sempower/>
- <https://yilinandrewang.shinyapps.io/pwrSEM/>
- <https://webpower.psychstat.org/wiki/models/index>

@@

- <https://schoemanna.shinyapps.io/mc_power_med/>
- <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>)
- <https://webpower.psychstat.org/wiki/models/index>

>>>>

The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
The pseudomedian can be computed using the `DescTools` package [@R-DescTools].
The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.

@@

A plot of statistical power is in @fig-nhst, as adapted from @Magnusson2013 (archived at <https://perma.cc/FG3J-85L6>).

@@

- <https://rpsychologist.com/pvalue> [@Magnusson2021; archived at <https://perma.cc/JP9F-9ZVY>]
- <https://rpsychologist.com/d3/pdist> [@Magnusson2015; archived at <https://perma.cc/BE96-8LSJ>]
- <https://rpsychologist.com/d3/nhst> [@Magnusson2014; archived at <https://perma.cc/ZU9A-37F3>]

@@

@JacksonWood2017 provides an online, interactive statistical decision tree to help you decide which statistical analysis to use: <https://www.statsflowchart.co.uk>

@@

Knowing any three of the following, you can calculate the fourth: statistical power, [effect size](#sec-practicalSignificance), [sample size](#sec-sampleVsPopulation), and [alpha level](#sec-statisticalSignificance).
Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
We use the `pwr` [@R-pwr] and `pwrss` [@R-pwrss] packages to estimate statistical power.
For free point-and-click software for calculating statistical power, see G\*Power: <https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower.html>

@@

- <https://aguinis.shinyapps.io/ml_power> [@Mathieu2012]
- <https://koumurayama.shinyapps.io/summary_statistics_based_power> [@Murayama2022]
- <https://webpower.psychstat.org/wiki/models/index> [@Zhang2018]

@@

- <https://sjak.shinyapps.io/power4SEM> [@Jak2020]
- <https://sempower.shinyapps.io/sempower> [@Moshagen2024]
- <https://yilinandrewang.shinyapps.io/pwrSEM> [@Wang2021]
- <https://webpower.psychstat.org/wiki/models/index> [@Zhang2018]

@@

- <https://schoemanna.shinyapps.io/mc_power_med> [@Schoemann2017]
- <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>) [@Ataneka2023], based on the `PowerUpR` package [@Bulus2021]
- <https://webpower.psychstat.org/wiki/models/index> [@Zhang2018]

### 2025-03-09 causal-inference.qmd
20250308 - citations

Another approach to control for variables is to use genetically informed designs.
Genetically informed designs allow controlling for potential genetic effects in order to more closely approximate the contributions of various environmental effects.
Genetically informed designs exploit differing degrees of genetic relatedness among participants to capture the extent to which genetic factors may contribute to an outcome.
The average percent of DNA shared between people of varying relationships is provided in @tbl-geneticRelatednessRelativePairs (<https://isogg.org/wiki/Autosomal_DNA_statistics>; archived at <https://perma.cc/MK3D-DST8>):

>>>>

Another approach to control for variables is to use genetically informed designs.
Genetically informed designs allow controlling for potential genetic effects in order to more closely approximate the contributions of various environmental effects.
Genetically informed designs exploit differing degrees of genetic relatedness among participants to capture the extent to which genetic factors may contribute to an outcome.
The average percent of DNA shared between people of varying relationships is provided by @ISOGG2022 in @tbl-geneticRelatednessRelativePairs (archived at <https://perma.cc/MK3D-DST8>):

### 2025-03-09 cognitive-bias.qmd
20250305 - hot hand citations
20250308 - citations

Consider the following example of the representativeness heuristic.
When coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.

@@

Here is a video of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:

@@

For example, the base rate of work-related injury is the general probability of experiencing a work-related injury, irrespective of other factors (e.g., the type of job, the person's age, the person's sex).
Among the working population in the U.S., the lifetime prevalence of work-related injuries (i.e., the percent of people who will experience a work-related injury at some point in their lives), is ~35% (<https://www.cdc.gov/mmwr/volumes/69/wr/mm6913a1.htm>; archived at <https://perma.cc/A2L6-WPEH>).
Thus, the base rate of work-related injuries in the U.S. is ~35%.
The probability of work-related injuries is higher for some occupations (e.g., construction) and for some groups (e.g., men, 55–64-year-olds, Black or Multiracial, who are self-employed and have less than high school education) than others.
Nevertheless, if we ignore all of the interacting factors, the general probability of work-related injuries is 35%.
If we made a prediction that someone would be highly likely (> 90%) to experience a work-related injury because they are male and self-employed, this would be ignoring the relatively lower base rate of work-related injury.
Indeed, even men (36.7%) and self-employed individuals (41.2%) have less than a 50% chance of experiencing a work-related injury.

@@

As applied to fantasy football, consider that you read about a potential sleeper Wide Receiver who had a stellar performance in a preseason game.
If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season (i.e., base rate neglect).
Performance in the preseason is not strongly predictive of performance in the regular season (<https://fivethirtyeight.com/features/the-nfl-preseason-is-not-predictive-but-it-can-often-seem-that-way/>; archived at <https://perma.cc/FSG2-6AXE>).

@@

An example of the regression fallacy is the so-called Sports Illustrated or Madden cover jinx curse.
The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated (the magazine) or Madden (the video game) will perform poorly.
But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>) [@Kahneman2011].
When a player has a superb season, they likely benefited from some degree to good luck, and it is unlikely that they will repeat such a stellar season the following year.
Instead, they are likely—at least based on random fluctuation—to regress to their long-term mean.

@@

Evidence on the hot hand is mixed.
Considerable evidence historically has suggested that there is no such thing as a "hot hand" even though many players and fans believe in it [@Gilovich1985; @BarEli2006; @Moskowitz2011; @Avugos2013; @Wetzels2016].
Even though streaks may occur, making a shot may not increase the likelihood of making the next shot; that is, the likelihood of making a field goal seems to be independent of past success (or failure) on recent attempts.
Defenses may adjust to guard a "hot" player more closely.
However, the findings suggesting no hot hand tend to hold even whether there is no defense [@Gilovich1985].
Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a].
However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].

>>>>

Consider the following example of the representativeness heuristic.
When coming out of college, Tight End Kyle Pitts drew comparisons to the "LeBron James" of Tight Ends [@Nivison2021; archived at <https://perma.cc/JQB5-XPVL>].
The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.

@@

Here is a video from @GetUpESPN2021 of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:

@@

For example, the base rate of work-related injury is the general probability of experiencing a work-related injury, irrespective of other factors (e.g., the type of job, the person's age, the person's sex).
Among the working population in the U.S., the lifetime prevalence of work-related injuries (i.e., the percent of people who will experience a work-related injury at some point in their lives), is ~35% [@Free2020; archived at <https://perma.cc/A2L6-WPEH>].
Thus, the base rate of work-related injuries in the U.S. is ~35%.
The probability of work-related injuries is higher for some occupations (e.g., construction) and for some groups (e.g., men, 55–64-year-olds, Black or Multiracial, who are self-employed and have less than high school education) than others.
Nevertheless, if we ignore all of the interacting factors, the general probability of work-related injuries is 35%.
If we made a prediction that someone would be highly likely (> 90%) to experience a work-related injury because they are male and self-employed, this would be ignoring the relatively lower base rate of work-related injury.
Indeed, even men (36.7%) and self-employed individuals (41.2%) have less than a 50% chance of experiencing a work-related injury.

@@

As applied to fantasy football, consider that you read about a potential sleeper Wide Receiver who had a stellar performance in a preseason game.
If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season (i.e., base rate neglect).
Performance in the preseason is not strongly predictive of performance in the regular season [@Schalter2022; archived at <https://perma.cc/FSG2-6AXE>].

@@

An example of the regression fallacy is the so-called Sports Illustrated or Madden cover jinx curse.
The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated (the magazine) or Madden (the video game) will perform poorly.
But, such a phenomenon can be more simply explained by regression to the mean [@Kahneman2011; @Smith2016a; archived at <https://perma.cc/CZM9-TVFN>].
When a player has a superb season, they likely benefited from some degree to good luck, and it is unlikely that they will repeat such a stellar season the following year.
Instead, they are likely—at least based on random fluctuation—to regress to their long-term mean.

@@

Evidence on the hot hand is mixed.
Considerable evidence historically has suggested that there is no such thing as a "hot hand" even though many players and fans believe in it [@Gilovich1985; @BarEli2006; @Moskowitz2011; @Avugos2013; @Wetzels2016].
Even though streaks may occur, making a shot may not increase the likelihood of making the next shot; that is, the likelihood of making a field goal seems to be independent of past success (or failure) on recent attempts.
Defenses may adjust to guard a "hot" player more closely.
However, the findings suggesting no hot hand tend to hold even whether there is no defense [@Gilovich1985].
Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a; @Miller2024].
However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].

### 2025-03-09 correlation.qmd
20250308 - R packages
20250308 - citations

Interactive visualizations by Kristoffer Magnusson on *p*-values and null-hypothesis significance testing are below:

- <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)

@@

>>>>

An interactive visualization by @Magnusson2020 on interpreting correlations is at the following link: <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)

@@

A nonlinear correlation can be estimated using the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:

### 2025-03-09 data-visualization.qmd
20250308 - citations

When designing graphics, it is important to understand general principles of graphic design.
An article that describes important principles about graphic design is at the following link: <https://www.adobe.com/express/learn/blog/8-basic-design-principles-to-help-you-create-better-graphics> (archived at <https://perma.cc/29P9-NNSK>).
The important principles include:

@@

The Data Visualization Catalogue provides examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>.
The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].
In this chapter, we will examine how to create statistical graphics to visualize data.
We will create the plots using the `ggplot2` package [@R-ggplot2; @Wickham2016_packages].
When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).

@@

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> (archived at <https://perma.cc/7VTA-Y8YS>)
- the `viridis` package [@R-viridis; @Garnier2024a_packages]
  - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- @Nunez2018

@@

For ways of visualizing univariate distributions, see [here](https://z3tt.github.io/beyond-bar-and-box-plots) (archived at <https://perma.cc/EEJ8-LND2>).

@@

Or, we could use a different theme, such as the dark theme (`theme_dark()`) in @fig-scatterplot10.
For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes>.

>>>>

When designing graphics, it is important to understand general principles of graphic design.
@AdobeExpress2020 describes important principles about graphic design at the following link: <https://www.adobe.com/express/learn/blog/8-basic-design-principles-to-help-you-create-better-graphics> (archived at <https://perma.cc/29P9-NNSK>).
The important principles include:

@@

The Data Visualization Catalogue provides examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>.
The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].
In this chapter, we will examine how to create statistical graphics to visualize data.
We will create the plots using the `ggplot2` package [@R-ggplot2; @ggplot3rdEd].
When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).

@@

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> [@Kilin2022; archived at <https://perma.cc/7VTA-Y8YS>]
- the `viridis` package [@R-viridis; @Garnier2024a_packages]
  - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- @Nunez2018

@@

@Scherer2021 describes various ways of visualizing univariate distributions; see [here](https://z3tt.github.io/beyond-bar-and-box-plots) (archived at <https://perma.cc/EEJ8-LND2>).

@@

Or, we could use a different theme, such as the dark theme (`theme_dark()`) in @fig-scatterplot10.
For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes> [@ggplot3rdEd].

### 2025-03-09 draft.qmd
20240830 - depth
20250227 - clean up
20250308 - citations

The Spurs in the National Basketball Association (NBA) were well-reputed for [excelling in this draft strategy](https://harvardsportsanalysis.org/2013/11/beating-the-nba-draft-does-any-team-outperform-expectations/) (archived at <https://perma.cc/X7NW-WZC6>).
They frequently used their second-round picks to draft high-risk, high-reward players.
Sometimes, the secound round pick was a bust, but they have little to lose with a failed second round pick.
Other times, their second round picks—including Willie Anderson, DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginóbili—greatly outperformed expectations.
Thanks, in part, to this draft strategy, the team showed strong extended success for nearly three decades from 1989 through the late-2010s.

@@

According to [an analysis](https://harvardsportsanalysis.wordpress.com/wp-content/uploads/2012/04/fantasyfootballdraftanalysis1.pdf) by the Harvard Sports Analysis Collective (archived at <https://perma.cc/P7RX-92UU>), the majority of the manager's salary cap should be spent on the starting lineup, and you should spend less on bench players.
This is known as the "stars and scrubs" draft strategy.
Based on the analysis, the author recommended applying a 10% premium to the top players and a 10% discount to the lower-tiered players.
The idea behind the approach is that a player on your bench does not contribute to the team's points and, thus, most players drafted to your bench do not contribute much to the team's points throughout the season.
That said, bench players can be important in the case of a starter's injury or under-performance.
So, it is recommended to draft starters with lower [uncertainty](#sec-fantasyValueUncertainty) who are safer.
In contrast to your starting lineup, you may look to draft players on your bench who have greater [uncertainty](#sec-fantasyValueUncertainty) for their high reward potential in a low-risk selection given the lower price.

An alternative to the "stars and scrubs" approach is to wait to draft more "high-value" players after other managers have over-paid for players.
In any case, having some small amount of cap left over toward the end of the draft can help you draft good value players to fill out your bench spots for cheap (e.g., $2).

>>>>

The Spurs in the National Basketball Association (NBA) were well-reputed for [excelling in this draft strategy](https://harvardsportsanalysis.org/2013/11/beating-the-nba-draft-does-any-team-outperform-expectations/) [@Ryan2013a; archived at <https://perma.cc/X7NW-WZC6>].
They frequently used their second-round picks to draft high-risk, high-reward players.
Sometimes, the secound round pick was a bust, but they have little to lose with a failed second round pick.
Other times, their second round picks—including Willie Anderson, DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginóbili—greatly outperformed expectations.
Thanks, in part, to this draft strategy, the team showed strong extended success for nearly three decades from 1989 through the late-2010s.

@@

According to [an analysis](https://harvardsportsanalysis.wordpress.com/wp-content/uploads/2012/04/fantasyfootballdraftanalysis1.pdf) by the Harvard Sports Analysis Collective, the majority of the manager's salary cap should be spent on the starting lineup, and you should spend less on bench players [@Chakravarthy2012; archived at <https://perma.cc/P7RX-92UU>].
This is known as the "stars and scrubs" draft strategy.
Based on the analysis, the author recommended applying a 10% premium to the top players and a 10% discount to the lower-tiered players.
The idea behind the approach is that a player on your bench does not contribute to the team's points and, thus, most players drafted to your bench do not contribute much to the team's points throughout the season.
That said, bench players can be important in the case of a starter's injury or under-performance.
So, it is recommended to draft starters with lower [uncertainty](#sec-fantasyValueUncertainty) who are safer.
In contrast to your starting lineup, you may look to draft players on your bench who have greater [uncertainty](#sec-fantasyValueUncertainty) for their high reward potential in a low-risk selection given the lower price.

An alternative to the "stars and scrubs" approach is to wait to draft more "high-value" players after other managers have over-paid for players.
In any case, having some small amount of cap left over toward the end of the draft can help you draft good value players to fill out your bench spots for cheap (e.g., $2).
Having some depth can help you offset the risk of injuries and the possibility that some of your starters may underperform their expectation, which is quite likely given how [challenging it is to predict fantasy performance](#sec-accuracyOfCrowd).

### 2025-03-09 evaluating-prediction-accuracy.qmd
20241105 - equation number
20250308 - R packages
20250308 - citations

- <http://araw.mede.uic.edu/cgi-bin/testcalc.pl>
- <https://dlrs.shinyapps.io/shinyDLRs>

@@

$$
\begin{aligned}
  \text{false omission rate (FOR)} &= P(C|\text{not } R) \\
  &= \frac{\text{FN}}{\text{FN} + \text{TN}} = 1 - \text{NPV}
\end{aligned}
$$ {#sec-falseOmissionRate}

@@

The adapted formula for MASE with non-time series data is described [here (https://stats.stackexchange.com/a/108963/20338)](https://stats.stackexchange.com/a/108963/20338) (archived at https://perma.cc/G469-8NAJ).
Scaled errors are calculated using @eq-scaledError:

@@

Predictive $R^2$ is described here: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>).
Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
Larger values indicate greater accuracy.

@@

The [area under the ROC curve (AUC)](#sec-auc) is a general index of discrimination accuracy for a categorical outcome.
It is also called the concordance ($c$) statistic.
Larger values reflect greater [discrimination accuracy](#sec-discrimination).
[AUC](#auc) was estimated using the `pROC` package [@R-pROC].

@@

As @Silver2012 notes, "The numbers have no way of speaking for themselves.
We speak for them.
We imbue them with meaning." (p. 9).
If we *understand* the variables in the system and how they influence each other, we can predict things more accurately than predicting for the sake of predicting.
For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) (archived at <https://perma.cc/PF8P-BT3D>), including extreme weather events like hurricanes.
These great strides have more to do with a better causal understanding of the weather system and the ability to conduct simulations of the atmosphere than merely because of big data [@Silver2012].
By contrast, other events are still incredibly difficult to predict, including earthquakes, in large part because we do not have a strong understanding of the system (and because we do not have ways of precisely measuring those causes because they occur at a depth below which we are realistically able to drill) [@Silver2012].

>>>>

- <http://araw.mede.uic.edu/cgi-bin/testcalc.pl> [@Schwartz2006; archived at <https://perma.cc/X8TF-7YBX>]
- <https://dlrs.shinyapps.io/shinyDLRs> [@Goodman2022]

@@

$$
\begin{aligned}
  \text{false omission rate (FOR)} &= P(C|\text{not } R) \\
  &= \frac{\text{FN}}{\text{FN} + \text{TN}} = 1 - \text{NPV}
\end{aligned}
$$ {#eq-falseOmissionRate}

@@

The adapted formula for MASE with non-time series data is described by @Hyndman2014 [here (https://stats.stackexchange.com/a/108963/20338)](https://stats.stackexchange.com/a/108963/20338) (archived at https://perma.cc/G469-8NAJ).
Scaled errors are calculated using @eq-scaledError:

@@

Predictive $R^2$ is described by @Hopper2014 here: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>).
Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
Larger values indicate greater accuracy.

@@

The [area under the ROC curve (AUC)](#sec-auc) is a general index of discrimination accuracy for a categorical outcome.
It is also called the concordance ($c$) statistic.
Larger values reflect greater [discrimination accuracy](#sec-discrimination).
[AUC](#auc) was estimated using the `pROC` package [@R-pROC; @Robin2011_packages].

@@

As @Silver2012 notes, "The numbers have no way of speaking for themselves.
We speak for them.
We imbue them with meaning." (p. 9).
If we *understand* the variables in the system and how they influence each other, we can predict things more accurately than predicting for the sake of predicting.
For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) [@Rosalsky2023; archived at <https://perma.cc/PF8P-BT3D>], including extreme weather events like hurricanes.
These great strides have more to do with a better causal understanding of the weather system and the ability to conduct simulations of the atmosphere than merely because of big data [@Silver2012].
By contrast, other events are still incredibly difficult to predict, including earthquakes, in large part because we do not have a strong understanding of the system (and because we do not have ways of precisely measuring those causes because they occur at a depth below which we are realistically able to drill) [@Silver2012].

### 2025-03-09 fantasy-football.qmd
20250308 - citations

Football is the most widely watched sport in the United States [archived at <https://perma.cc/X2UG-RAAK>; @Jones2024; archived at <https://perma.cc/JNU6-S96A>; @statistica2023].

@@

Here is a video that provides a brief introduction to fantasy football:

>>>>

Football is the most widely watched sport in the United States [@Jones2024; archived at <https://perma.cc/X2UG-RAAK>; @statistica2023; archived at <https://perma.cc/JNU6-S96A>].

@@

Here is a video that provides a brief introduction to fantasy football [@NFLFantasy2020]:

### 2025-03-09 index.qmd
20250308 - citations

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics", "pROC", "rms"
)

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
  "nflreadr", "nflplotR", "nflfastR", "nfl4th", "gsisdecoder", "progressr",
  "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag",
  "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC",
  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg",
  "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis",
  "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
  "caret", "gpboost", "cmdstanr", "learnr", "remotes" #"mgcv",
)
